\documentclass[a4paper]{scrartcl}

\usepackage{minted}

\usepackage{tikz}
\usetikzlibrary{automata, positioning, arrows}
\tikzset{
	->, % makes the edges directed
	>=stealth', % makes the arrow heads bold
	node distance=3cm, % specifies the minimum distance between two nodes. Change if necessary.
	every state/.style={thick, fill=gray!10}, % sets the properties for each ’state’ node
	initial text=$ $, % sets the text that appears on the start arrow
}

\usepackage{amsthm, amssymb, amsmath}
\usepackage{bm}

\newtheorem{theorem}{Theorem}
\newtheorem*{theorem*}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem*{lemma*}{Lemma}

\newtheorem{claim}{Claim}
\newtheorem*{claim*}{Claim}

\theoremstyle{definition}
\newtheorem{definition}{Definition}
\newtheorem*{definition*}{Definition}

\newcommand{\eps}{\varepsilon}
\newcommand{\card}[1]{\left\lvert #1 \right\rvert}

\newcommand{\NN}{\mathbb N}

\title{
	Programming Languages: Lecture 7\\
	Regex
}
\author{Rishabh Dhiman}
\date{15 January 2022}

\begin{document}
\maketitle

\section{Deterministic Finite Automaton}
For a deterministic automaton,
\begin{enumerate}
	\item No $\eps$-transitions exist
	\item Exactly one transition exists for every character and state.
\end{enumerate}
We can convert NFA to DFA using the powerset construction.

\section{Scanning with Output}
Scanner is very similar to a scanner.
\begin{itemize}
	\item DFA just accepts or rejects a token/lexeme. Scanner needs to classify them all.
	\item Algorithms minimize number of accepting states of DFA, this is not desirable in scanners where you need to classify the strings.
\end{itemize}

\section{Production rules}
Form production rules from the DFA.
\begin{align*}
	S &\to aA \mid bB\\
	A &\to aA \mid bC\\
	B &\to aA \mid bB\\
	C &\to aA \mid bD\\
	D &\to aA \mid bB \mid \eps
\end{align*}
You start with the start state, $S$ here. Then apply the production rules to generate strings. End at a terminal variable, one with a transition to $\eps$.

This is a regular or right-linear grammar as all the variables come at the end in the production rules.

We can have production rules which don't generate a regular language. Consider the language
\[P = \{a^n b^n \mid n \in \NN_0\}.\]
$P$ is not regular (there exists $i \neq j$ such that $a^i$ and $a^j$ end at same state in DFA, which, $a^i b^i$ being accepted implies $a^j b^i$ gets accepted) but can easily be represented by the production rule
\[X \to a X b \mid \eps.\]

\section{Grammar}
\begin{definition}[Grammar]
A grammar $G = \langle N, T, P, S\rangle$ consists of
\begin{itemize}
	\item a set $N$ of nonterminal symbols,
	\item a start symbol $S \in N$,
	\item a set $T$ of terminal symbols or the alphabet,
	\item a set $P$ of productions or rewrite rules where each rule is of the form $\alpha \to \beta$ for $\alpha, \beta \in (N \cup T)^*$.
\end{itemize}
\end{definition}
\begin{definition}
	Given a grammar $G$, any $\alpha \in (N\cup T)^*$ is called a sentential form. Any $x \in T^*$ is called a sentence.
\end{definition}
\end{document}
